Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-09 inc/photocathode/este/este.tex
merge most of the remaining comments from andreas on the PAPE section

\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
Prior studies \cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included \cite{dowell_quantum_2009}.
Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
Note that due to the surprisingly large yield, we repeated the measurements, this time replacing the the scintillator screen with a Faraday cup and repeating the laser energy sequence.
In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in Sections \ref{sec:single_photon} and \ref{sec:two_photon_thermionic}, for which relative measurements were sufficient.

>>>>

\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
Prior studies \cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included \cite{dowell_quantum_2009}.
Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
Note that due to the surprisingly large yield, we repeated the measurements, this time replacing the the scintillator screen with a Faraday cup and repeating the laser energy sequence.
In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in the preceeding sections, for which relative measurements were sufficient.

### 2013-04-09 inc/photocathode/plasmon/plasmon.tex
merge most of the remaining comments from andreas on the PAPE section

% Content taken from 2010 ANL report
Note that for the following measurement, relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
The scintillator and CCD camera detection scheme was also shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
As a result, the detection scheme will accurately determine any laser power law dependence to the electron emission from the photocathode.

Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
These two properties are known characteristics of plasmon-assisted photoemission.

As seen in \ref{fig:pape}, at low incident laser intensities, below about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
Moreover, both emissions have similar momentum distributions (\ref{fig:pape} A and E), again indicating a similar emission process.
The only difference is that the emission from the nano-patterned region is 3-4 times more efficient.
This is readily explained by the fact that the periodic surface modulation results in a local enhancement of the $\sim$1MV/m DC gun field at the peaks of the modulation which, through the Schottky effect, then presents a reduced barrier for field emission.

Above $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
The observed distinct changes in the momentum distributions (\ref{fig:pape} B and C) in this laser intensity range support this argument.
In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves). 
This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \ref{fig:pape} D and by the cut-off in the detected electron signal (YAG fluorescence).

@@

In principle this unexpected increase in transverse momentum could be caused by any of three effects.
The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
The possibility is easily dismissed as the effect is not observed when off the grating.
The second possible cause is the divergence of the plasmon's own electric field ($\partial E_{plasmon} / \partial x$).
This field is parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
The third possible cause is local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface, commonly known as the ``lightning rod'' effect.
For the employed one dimensional grating, this effect is necessarily parallel to the grating vector $k_G$.

\begin{figure}
  \centering
  \input{rotated}
  \caption[Comparison of the colinear and rotated plasmonic excitation geometries]{
    Comparison of the colinear (above) and rotated (below) plasmonic excitation geometries.
    This rotation separates the directionality of effects caused by the laser, plasmon field and grating.
    Each vector diagrams (left) is shown next to a representative observed momentum distribution image (right).
    The increased divergence is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed); evidence that the grating is the cause of the deleterious effect.
  }
  \label{fig:rotated}
\end{figure}

Given the geometry of the previous measurement, notably the colinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.
\ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
For this rotation, the laser angle must be increased from $39^{\circ}$ to $47^{\circ}$.

The results are shown to the right in \ref{fig:rotated}.
As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
As that geometry does not employ a patterned surface, it would is not subject to such local field enhancements of the accelerating electric field.

@@

Though the cause of the failure may never be known for certain, several causes are suspected.
Now that the surface is not purely sinusoidal, the surface has multiple Fourier components.
This may mean that the laser may not couple as efficiently to the plasmon, requiring increased laser power for oscillation. 
Additionally, the non-smooth surface may cause increaded difficulty for propagating the plasmon for large distances.
Finally, upon further investigation of the surface by SEM, evidence seen of incomplete filling of the gold into the trenches, which may have interrupted creation or propagation of the plasmon.

@@

Although plasmonic photocathodes are still an interesting field of research in photocathode physics, we have concluded that they are not a promising candidate for an electron source for UEM at this time.
They have been difficult to manufacture and thus would be expensive for owners.
They are very intolerant to misalignment and would therefore be problematic for users.
Finally, they cannot be shown to produce the performance (that is, a reduced transverse emittance) needed to justify the previous concerns.

Other nano-engineered surfaces, including quantum-wells, are still being considered for future work.

>>>>

% Content taken from 2010 ANL report
\ref{fig:pape} summarizes the photoemission results obtained from this gold-coated periodic photocathode under green (523nm) laser irradiation.
Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
These two properties are known characteristics of plasmon-assisted photoemission \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.

As seen in \ref{fig:pape}, at low incident laser intensities, below a laser intensity of about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
Moreover, both emissions have similar momentum distributions (\ref{fig:pape} A and E), again indicating a similar emission process.
The only difference is that the emission from the nano-patterned region is 3-4 times more efficient.
This is readily explained by the fact that the periodic surface modulation results in a local enhancement of the $\sim$1MV/m DC gun field at the peaks of the modulation which, through the Schottky effect, then presents a reduced barrier for field emission.

I note here that for these measurements, the relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
The scintillator and CCD camera detection scheme was shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
%TODO introduce this linearity earlier?
As a result, the detection scheme will accurately determine any laser power law dependence to the electron emission from the photocathode.

Above an incident laser pulse intensity of $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
The observed distinct changes in the momentum distributions (\ref{fig:pape} B and C) in this laser intensity range support this argument.
In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves). 
This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \ref{fig:pape} D and by the cut-off in the detected electron signal (YAG fluorescence).

@@

In principle, this unexpected increase in transverse momentum could be caused by any of three effects.
The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
The possibility is easily dismissed as the effect is not observed when off the grating.
The second possible cause is the divergence of the plasmon's own electric field ($\partial E_{plasmon} / \partial x$).
This field is parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
The third possible cause is local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface, commonly known as the ``lightning rod'' effect.
For the employed one dimensional grating, this effect is necessarily parallel to the grating vector $k_G$.

\begin{figure}
  \centering
  \input{rotated}
  \caption[Comparison of the colinear and rotated plasmonic excitation geometries]{
    Comparison of the collinear (above) and rotated (below) plasmonic excitation geometries.
    This rotation separates the directionality of effects caused by the laser, plasmon field and grating.
    Each vector diagrams (left) is shown next to a representative observed momentum distribution image (right).
    The increased divergence is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed); evidence that the grating is the cause of the deleterious effect.
  }
  \label{fig:rotated}
\end{figure}

Given the geometry of the previous measurement, notably the collinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.
The left side of \ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
For this rotation, the laser angle must be increased from $39^{\circ}$ to $47^{\circ}$.

The experimental results are shown to the right in \ref{fig:rotated}.
As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
As that geometry does not employ a patterned surface, it would not be subject to such local field enhancements of the accelerating electric field.

@@

Though the cause of the failure may never be known for certain, several causes are suspected.
Now that the surface is not purely sinusoidal, so that the surface has multiple Fourier components.
This may mean that the laser may not couple as efficiently to the plasmon, requiring increased laser power for oscillation. 
Additionally, the non-smooth surface may cause increased difficulty for propagating the plasmon for large distances.
Finally, upon further investigation of the surface by SEM, evidence was observed of incomplete filling of the gold coating into the trenches, which may have interrupted creation or propagation of the plasmon.

@@

Although plasmonic photocathodes are still an interesting field of research in photocathode physics, we have concluded that they are not a promising candidate for an electron source for UEM at this time.
They have been difficult to manufacture and thus would likely be expensive.
They are very intolerant to misalignment and would therefore be problematic for users.
Finally, they cannot be shown to produce the performance (that is, a reduced transverse emittance) needed to justify the previous concerns.

Other nano-engineered surfaces are still being considered for future work.
Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013}.
Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon \cite{kupersztych_anomalous_2005}.
Photoemission from nano-arrays of quantum-wells also offers the tantalizing hope of reduced rms transverse momentum of the generated electrons.
